---
title: "12_DataVisualization.rmd"
output: html_document
date: "2025-03-13"
---
```{r}
source("00_requirements.R")
source("02_func_dVisual_PlotDensity.R")
df <- read.csv("cleaned_sleep_health_data.csv")
```

1. test correlation of sleep quality and stress level
```{r, warning = FALSE}
# check whether two variables are normally distributed. If yes, use Pearson, if no,  use Spearman’s
shapiro_sleep <- shapiro.test(df$sleep_quality)
shapiro_stress <- shapiro.test(df$stress)

# p-value to determine its normality
if (shapiro_sleep$p.value > 0.05 & shapiro_stress$p.value > 0.05) {
  # Pearson correlation
  correlation_test <- cor.test(df$sleep_quality, df$stress, method = "pearson")
} else {
  # Spearman correlation
  correlation_test <- cor.test(df$sleep_quality, df$stress, method = "spearman")
}
print(correlation_test)

# Visualization of correlation
ggscatter(df, x = "stress", y = "sleep_quality",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = correlation_test$method,
          xlab = "Stress Level", ylab = "Sleep Quality")
```
- P-value: 2.2e-16
- rho value:  -0.908
- negative trend line on the scatter plot
combining the three information, the two variables, sleep quality and stress level, have a high correlation. (more analysis in the final report)



2. Conduct t-tests on 6 hypothesis. Comparing the difference of two means.
```{r}
# 1. sleep quality of sleep apnea vs sleep quality of insomnia
t_test_result1 <- t.test(df$sleep_quality[df$disorder == "Sleep Apnea"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df,"sleep_quality", "Sleep Apnea", "Insomnia", "Figure1: Sleep Quality: Sleep Apnea vs. Insomnia")
t_test_result1
```
- p value = 0.001583 < 0.05:
        we reject the null hypothesis, indicating the difference in sleep quality between individuals with Sleep Apnea and Insomnia is statistically significant.
- means: 7.205128(Sleep Apnea) vs 6.532468(Insomnia):
        Individuals with Sleep Apnea tend to report better sleep quality than those with Insomnia.
- 95 percent confidence interval: 0.2610652 1.0842562:
        The range does not contain 0, further confirming that the difference in means is statistically significant.

```{r}
# 2. the stress level with the ppl with sleep apnea vs thestress level with the ppl with insomnia \
t_test_result2 <- t.test(df$stress[df$disorder == "Sleep Apnea"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "Sleep Apnea", "Insomnia", "Figure2: Stress Level: Sleep Apnea vs. Insomnia")
t_test_result2

```
- p-value = 0.5161 > 0.05:
        We fail to reject the null hypothesis, indicating that the difference in stress levels between individuals with Sleep Apnea and Insomnia is not statistically significant.
- Means: 5.666667 (Sleep Apnea) vs. 5.870130 (Insomnia):
        The average stress levels for both groups are very close, suggesting no significant difference in stress between these two sleep disorders.
- 95 percent confidence interval: (-0.8216114, 0.4146850):
        The interval contains 0, confirming that the difference in means is not statistically significant.

```{r}
# 3. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with insomnia 

t_test_result3 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df, "sleep_quality", "None", "Insomnia", "Figure3: Sleep Quality: No Disorder vs. Insomnia")
t_test_result3
```


```{r}
# 4. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with sleep apnea
t_test_result4 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Sleep Apnea"])
plot_density(df, "sleep_quality", "None", "Sleep Apnea", "Figure4: Sleep Quality: No Disorder vs. Sleep Apnea")
t_test_result4
# p value: 0.03601
# means: 7.625571  7.205128
# 95 percent confidence interval: 0.02801647 0.81286868
```
```{r}
# 5. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with insomnia 
t_test_result5 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "None", "Insomnia", "Figure5: Stress Level: No Disorder vs. Insomnia")
t_test_result5
```
p value: 0.0002053
means: 5.114155  5.870130
95 percent confidence interval: -1.1481604 -0.3637889

analysis:

```{r}
# 6. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with sleep apnea. 
t_test_result6 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Sleep Apnea"])
plot_density(df, "stress", "None", "Sleep Apnea", "Figure6: Stress Level: No Disorder vs. Sleep Apnea")
t_test_result6
```
p value: 0.0555
means: 5.114155  5.666667
95 percent confidence interval: -1.11825552  0.01323269

analysis: