---
title: "12_DataVisualization.rmd"
output: html_document
date: "2025-03-13"
---
```{r}
source("00_requirements.R")
source("02_func_dVisual_PlotDensity.R")
df <- read.csv("cleaned_sleep_health_data.csv")
```

```{r}
# 1. sleep quality of sleep apnea vs sleep quality of insomnia
t_test_result1 <- t.test(df$sleep_quality[df$disorder == "Sleep Apnea"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df,"sleep_quality", "Sleep Apnea", "Insomnia", "Sleep Quality: Sleep Apnea vs. Insomnia")
t_test_result1
```

```{r}
# 2. the stress level with the ppl with sleep apnea vs thestress level with the ppl with insomnia \
t_test_result2 <- t.test(df$stress[df$disorder == "Sleep Apnea"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "Sleep Apnea", "Insomnia", "Stress Level: Sleep Apnea vs. Insomnia")
t_test_result2
```

```{r}
# 3. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with insomnia 

t_test_result3 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df, "sleep_quality", "None", "Insomnia", "Sleep Quality: No Disorder vs. Insomnia")
t_test_result3
```

```{r}
# 4. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with sleep apnea
t_test_result4 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Sleep Apnea"])
plot_density(df, "sleep_quality", "None", "Sleep Apnea", "Sleep Quality: No Disorder vs. Sleep Apnea")
t_test_result4
```

```{r}
# 5. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with insomnia 
t_test_result5 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "None", "Insomnia", "Stress Level: No Disorder vs. Insomnia")
t_test_result5
```

```{r}
# 6. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with sleep apnea. 
t_test_result6 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Sleep Apnea"])
plot_density(df, "stress", "None", "Sleep Apnea", "Stress Level: No Disorder vs. Sleep Apnea")

cor.test(df$sleep_quality, df$stress, method = "pearson")
t_test_result6
```

## stats107_final


##How do different sleep disorders (Insomnia vs. Sleep Apnea) differ in sleep quality and stress level?

# 1. Variables:

  - Sleep Disorder: The presence or absence of a sleep disorder in the person (None, Insomnia, Sleep Apnea).

  - Quality of Sleep (Scale: 1-10): A subjective rating of the quality of sleep, ranging from 1 to 10.

  - Stress Level (Scale: 1-10): A subjective rating of the stress level experienced by the person, ranging from 1 to 10.


# 2. Hypotheses:
All the following tests will use a t-test to compare the differences of the two means.


  1) Null Hypothesis (H₀₁) : sleep quality of sleep apnea vs sleep quality of insomnia
  
      - There is no significant difference in sleep quality between people with Sleep Apnea and those with Insomnia.

  2) Null Hypothesis (H₀₂) : stress level of sleep apnea vs stress level of insomnia

      - There is no significant difference in stress levels between people with Sleep Apnea and those with Insomnia.

  3) Null Hypothesis (H₀₃) : sleep quality of normal people vs sleep quality of insomnia

     - There is no significant difference in sleep quality between people with No Sleep Disorder and those with Insomnia.

  4) Null Hypothesis (H₀₄) : sleep quality of normal people vs sleep quality of sleep apnea

     - There is no significant difference in sleep quality between people with No Sleep Disorder and those with Sleep Apnea.

  5) Null Hypothesis (H₀₅) : stress level of normal people vs stress level of insomnia
  
     - There is no significant difference in stress levels between people with No Sleep Disorder and those with Insomnia.
  
  6) Null Hypothesis (H₀₆) : stress level of normal people vs stress level of sleep apnea
  
     - There is no significant difference in stress levels between people with No Sleep Disorder and those with Sleep Apnea.
