---
title: "12_DataVisualization.rmd"
output: html_document
date: "2025-03-13"
---
```{r}
source("00_requirements.R")
source("02_func_dVisual_PlotDensity.R")
df <- read.csv("cleaned_sleep_health_data.csv")
```


```{r, warning = FALSE}
#I want to check whether sleep quality and stress level are correlated

# check whether it's normally distributed. If yes, use Pearson, if no,  use Spearman’s
shapiro_sleep <- shapiro.test(df$sleep_quality)
shapiro_stress <- shapiro.test(df$stress)

# p-value to determine its normality
if (shapiro_sleep$p.value > 0.05 & shapiro_stress$p.value > 0.05) {
  # Pearson correlation
  correlation_test <- cor.test(df$sleep_quality, df$stress, method = "pearson")
} else {
  # Spearman correlation
  correlation_test <- cor.test(df$sleep_quality, df$stress, method = "spearman")
}
print(correlation_test)

# Visualization of correlation
ggscatter(df, x = "stress", y = "sleep_quality",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = correlation_test$method,
          xlab = "Stress Level", ylab = "Sleep Quality")
```
P-value <- 2.2e-16
The p-value is extremely small, meaning the correlation is highly statistically significant. This suggests that the relationship between stress and sleep quality is not due to random chance. The high correlation also being confirmed by the -0.908 rho value and the scatter plot, indicates a strong negative correlation between sleep quality and stress. 

```{r}
# 1. sleep quality of sleep apnea vs sleep quality of insomnia
t_test_result1 <- t.test(df$sleep_quality[df$disorder == "Sleep Apnea"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df,"sleep_quality", "Sleep Apnea", "Insomnia", "Figure1: Sleep Quality: Sleep Apnea vs. Insomnia")
t_test_result1
```
p value: 0.001583
means: 7.205128  6.532468
95 percent confidence interval: 0.2610652 1.0842562

analysis: 
```{r}
# 2. the stress level with the ppl with sleep apnea vs thestress level with the ppl with insomnia \
t_test_result2 <- t.test(df$stress[df$disorder == "Sleep Apnea"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "Sleep Apnea", "Insomnia", "Figure2: Stress Level: Sleep Apnea vs. Insomnia")
t_test_result2

```
p value: 0.5161
means: 5.666667  5.870130
95 percent confidence interval: -0.8216114  0.4146850

analysis: 

```{r}
# 3. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with insomnia 

t_test_result3 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df, "sleep_quality", "None", "Insomnia", "Figure3: Sleep Quality: No Disorder vs. Insomnia")
t_test_result3
```

```{r}
# 4. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with sleep apnea
t_test_result4 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Sleep Apnea"])
plot_density(df, "sleep_quality", "None", "Sleep Apnea", "Figure4: Sleep Quality: No Disorder vs. Sleep Apnea")
t_test_result4
# p value: 0.03601
# means: 7.625571  7.205128
# 95 percent confidence interval: 0.02801647 0.81286868
```
p value: 2.2e-16
means: 7.625571  6.532468
95 percent confidence interval: 0.8701568 1.3160497

analysis: 
```{r}
# 5. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with insomnia 
t_test_result5 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "None", "Insomnia", "Figure5: Stress Level: No Disorder vs. Insomnia")
t_test_result5
```
p value: 0.0002053
means: 5.114155  5.870130
95 percent confidence interval: -1.1481604 -0.3637889

analysis:

```{r}
# 6. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with sleep apnea. 
t_test_result6 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Sleep Apnea"])
plot_density(df, "stress", "None", "Sleep Apnea", "Figure6: Stress Level: No Disorder vs. Sleep Apnea")
t_test_result6
```
p value: 0.0555
means: 5.114155  5.666667
95 percent confidence interval: -1.11825552  0.01323269

analysis: