---
title: "12_DataVisualization.rmd"
output: html_document
date: "2025-03-13"
---
```{r}
source("00_requirements.R")
source("02_func_dVisual_PlotDensity.R")
df <- read.csv("cleaned_sleep_health_data.csv")
```

1. test correlation of sleep quality and stress level
```{r, warning = FALSE}
# check whether two variables are normally distributed. If yes, use Pearson, if no,  use Spearman’s
shapiro_sleep <- shapiro.test(df$sleep_quality)
shapiro_stress <- shapiro.test(df$stress)

# p-value to determine its normality
if (shapiro_sleep$p.value > 0.05 & shapiro_stress$p.value > 0.05) {
  # Pearson correlation
  correlation_test <- cor.test(df$sleep_quality, df$stress, method = "pearson")
} else {
  # Spearman correlation
  correlation_test <- cor.test(df$sleep_quality, df$stress, method = "spearman")
}
print(correlation_test)

# Visualization of correlation
ggscatter(df, x = "stress", y = "sleep_quality",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = correlation_test$method,
          xlab = "Stress Level", ylab = "Sleep Quality",
          main = "Correlation of Sleep Quality and Stress Level")
```
- P-value: 2.2e-16
- rho value:  -0.908
- negative trend line on the scatter plot
combining the three information, we conclude that sleep quality and stress level have a high correlation. Thus, we can research on how different they are in individuals, specifically with different sleep disorder and without any sleep disorder. (more analysis in the final report)



2. Conduct t-tests on 6 hypothesis. Comparing the difference of two means.
(Note: due to the formatting for the final report, the size of the font and the plots are abnormal here.)
```{r}
# 1. sleep quality of sleep apnea vs sleep quality of insomnia
t_test_result1 <- t.test(df$sleep_quality[df$disorder == "Sleep Apnea"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df,"sleep_quality", "Sleep Apnea", "Insomnia", "Figure1: Sleep Quality: Sleep Apnea vs. Insomnia")
t_test_result1
```
- p value = 0.001583 < 0.05:
        we reject the null hypothesis, indicating the difference in sleep quality between individuals with Sleep Apnea and Insomnia is statistically significant.
- means: 7.205128(Sleep Apnea) vs 6.532468(Insomnia):
        Individuals with Sleep Apnea tend to report better sleep quality than those with Insomnia.
- 95 percent confidence interval: 0.2610652 1.0842562:
        The range does not contain 0, further confirming that the difference in means is statistically significant.

```{r}
# 2. the stress level with the ppl with sleep apnea vs thestress level with the ppl with insomnia \
t_test_result2 <- t.test(df$stress[df$disorder == "Sleep Apnea"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "Sleep Apnea", "Insomnia", "Figure2: Stress Level: Sleep Apnea vs. Insomnia")
t_test_result2

```
- p-value = 0.5161 > 0.05:
        We fail to reject the null hypothesis, indicating that the difference in stress levels between individuals with Sleep Apnea and Insomnia is not statistically significant.
- Means: 5.666667 (Sleep Apnea) vs. 5.870130 (Insomnia):
        The average stress levels for both groups are very close, suggesting no significant difference in stress between these two sleep disorders.
- 95 percent confidence interval: (-0.8216114, 0.4146850):
        The interval contains 0, confirming that the difference in means is not statistically significant.

```{r}
# 3. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with insomnia 
t_test_result3 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Insomnia"])
plot_density(df, "sleep_quality", "None", "Insomnia", "Figure3: Sleep Quality: No Disorder vs. Insomnia")
t_test_result3
```
- p-value < 2.2e-16:
        We reject the null hypothesis, indicating a significant difference in sleep quality between individuals with no sleep disorder and those with Insomnia.
- Means: 7.625571 (None) vs. 6.532468 (Insomnia):
        Individuals without sleep disorders report significantly better sleep quality than those with Insomnia.
- 95% Confidence Interval: (0.8701568, 1.3160497):
The interval does not contain 0, further confirming the significant difference.


```{r}
# 4. the quality of sleeping with the ppl without sleeping disorder vs the quality of sleeping with the ppl with sleep apnea
t_test_result4 <- t.test(df$sleep_quality[df$disorder == "None"],
       df$sleep_quality[df$disorder == "Sleep Apnea"])
plot_density(df, "sleep_quality", "None", "Sleep Apnea", "Figure4: Sleep Quality: No Disorder vs. Sleep Apnea")
t_test_result4
```
- p-value = 0.03601 < 0.05:
        We reject the null hypothesis, indicating that the difference in sleep quality between individuals with no disorder and those with Sleep Apnea is statistically significant.
- Means: 7.625571 (None) vs. 7.205128 (Sleep Apnea):
        Individuals without sleep disorders report slightly better sleep quality than those with Sleep Apnea.
- 95% Confidence Interval: (0.02801647, 0.81286868):
        The interval does not contain 0, confirming the statistical significance of the difference.
```{r}
# 5. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with insomnia 
t_test_result5 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Insomnia"])
plot_density(df, "stress", "None", "Insomnia", "Figure5: Stress Level: No Disorder vs. Insomnia")
t_test_result5
```
- p-value = 0.0002053 < 0.05:
        We reject the null hypothesis, indicating a significant difference in stress levels between individuals with no sleep disorder and those with Insomnia.
- Means: 5.114155 (None) vs. 5.870130 (Insomnia):
        Individuals with Insomnia report higher stress levels compared to those with no sleep disorder.
- 95% Confidence Interval: (-1.1481604, -0.3637889):
        The confidence interval does not contain 0, confirming the statistical significance of the difference.

```{r}
# 6. the stress level with the ppl without sleeping disorder vs the stress levelwith the ppl with sleep apnea. 
t_test_result6 <- t.test(df$stress[df$disorder == "None"],
       df$stress[df$disorder == "Sleep Apnea"])
plot_density(df, "stress", "None", "Sleep Apnea", "Figure6: Stress Level: No Disorder vs. Sleep Apnea")
t_test_result6
```

- p-value = 0.0555 > 0.05:
        We fail to reject the null hypothesis, suggesting that there is no statistically significant difference in stress levels between individuals with no sleep disorder and those with Sleep Apnea.
- Means: 5.666667 (Sleep Apnea) vs. 5.870130 (Insomnia):
        The average stress levels for both groups are very close, suggesting no significant difference in stress between these two sleep disorders.
  - 95% Confidence Interval: (-1.11825552,  0.01323269):
        Since the confidence interval contains 0, it confirms the lack of statistical significance.