---
title: "13_data_visualization"
output: pdf_document
date: "2025-03-15"
---

```{r}
source("00_requirements.R")
source("01_func_DataCleaning.R")
source("03_func_dVisual_BoxPlot.R")
df <- read.csv("cleaned_sleep_health_data.csv")
```
After conducting the **t-test** and visualizing the **distribution of sleep quality and stress levels** with density plots, we further analyze the data using **box plots**. 
The box plots provide a visual comparison of sleep quality and stress levels across different groups based on sleep disorder status (sleep apnea, insomnia, none). By examining the distribution of values within each group, we can observe key differences in central tendency and variability.

```{r}
# Plot sleep quality for all three groups
plot_boxplot(df, "sleep_quality", "Figure7: Sleep Quality Across Sleep Disorders")

# Plot stress level for all three groups
plot_boxplot(df, "stress", "Figure8: Stress Levels Across Sleep Disorders")
```


```{r}
ggplot(df, aes(x = stress, y = sleep_quality, shape = disorder, color = disorder)) +  
  # Create a scatter plot with:
  # - X-axis: Stress level
  # - Y-axis: Sleep quality
  # - Shape & color: Coded by 'disorder' category
  
  geom_point(size = 3, alpha = 0.7) +  
  # Plot individual data points with:
  # - Size = 3 (moderate point size)
  # - Alpha = 0.7 (70% opacity for better visualization)
  
  geom_hline(yintercept = median(df$sleep_quality, na.rm = TRUE), linetype = "dashed", color = "black") +  
  # Add a horizontal dashed line at the median sleep quality to divide the plot into quadrants
  
  geom_vline(xintercept = median(df$stress, na.rm = TRUE), linetype = "dashed", color = "black") +  
  # Add a vertical dashed line at the median stress level to complete the four-quadrant separation
  
  labs(
    title = "Four-Quadrant Distribution of Stress & Sleep Quality",  # Set the plot title
    x = "Stress Level",  # Label for the X-axis
    y = "Sleep Quality",  # Label for the Y-axis
    shape = "Disorder",  # Label for shape legend
    color = "Disorder"   # Label for color legend
  ) +
  
  scale_shape_manual(values = c(16, 17, 15)) +  
  # Assign specific shapes for different disorder categories:
  # - 16 (solid circle) for one group
  # - 17 (triangle) for another
  # - 15 (square) for the third
  
  theme_minimal()  
  # Apply a minimalistic theme for a clean, professional appearance

```
